---
title: "AstraZenica SouthAfrica"
author: "Dr K"
date: "2/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tigerstats)
require(tidyverse)
require(knitr)
```

## Create a table from the published data


```{r}
groupsize = 2000
subgp = groupsize/2
vaccine = c(subgp-19,19)
placebo =  c(subgp-23,23)
results = as.table(cbind(vaccine,placebo))
rownames(results) <- c("none","covid")

```

## Look at a "percentage barchart" of the reported cases.  If the vaccine is innefective then covid should be more or less equally distributed between the vaccine and the placebo.  The more effective the vaccine the lower would be the percent in the group of covid.

```{r}
barchartGC(results,type = "percent")
```
  
## Hmmm, that barchart shows there is not much difference between the vaccine and the placebo.  Important to note this is the B.1.351 virus variant

```{r}
fisher.test(results)
```

## The fisher exact test is an odds-ratio test. It is commonly used in clinical trials to test the effectiveness of preventitive drugs like vaccines. In this case it says the odds of getting covid is 1.215 times greater if you had taken the placebo than if you had taken the vaccine. The media is reporting this as hardly effective. It also gives you a confidence interval that has the true odds ratio within it 95% of the time.  That interval is .62 to 2.38

```{r}
effectiveness =  1.215  /( 1.215   +1)
print(effectiveness*100)
```

```{r}
low=(.62/(.62 + 1))*100
hi=(2.38/(2.38 +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`