---
title: "Sputnick Vaccine Russian"
author: "Dr K"
date: "2/2/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tigerstats)
require(tidyverse)
require(knitr)
```

# What they did  

## The preliminary findings were published in The Lancet on Tuesday (Sep 2, 2021) and are based on data gathered from 19,866 participants, of which around three-quarters (14,964) received two doses of the vaccine and a quarter (4,902) were given two doses of a placebo. Sixteen cases of symptomatic Covid-19 were confirmed in the vaccine and sixty two cases were found in the placebo group -- equating to an efficacy of 91.6% (according to their analysis).

Most Clinical trials have equal sized groups - but Russia made a more sensible approach by having a larger vaccine group (betting on the come so-to-speak).  This is a good one for beginning statistics students so they learn their analysis skills on real data.

```{r}

vaccine = c(14964-16,16)
placebo =  c(4902-62,62)
results = as.table(cbind(vaccine,placebo))
rownames(results) <- c("none","covid")

```

## We show two barcharts - The first is results of the study and the second is what it would look like if the vaccine was not effective.



```{r}
barchartGC(results,type = "percent", main="Results from the Sputnik V Clinical Trial")
```
  
We run a chi-sq test using Chisq Gc in order to obtain expected results when vaccine does not work.

```{r}
chisqtestGC(results)
vaccine2=c(14905,59)
placebo2=c(4882,19)
resultsExp = as.table(cbind(vaccine2,placebo2))
rownames(resultsExp) <- c("none","covid")
barchartGC(resultsExp,type = "percent",main="Expected results of ineffective vaccine \n With the same patient assignment structure")
```
  
## Notice how different the two barcharts appear.  Even your grandma could see that most of the people that got covid in the Sputnik V trial came from the group of people who received the placebo!

```{r}

fisher.test(results)
fisher.test(resultsExp)
```
```{r}
fish <- fisher.test(results)
options(digits=2)
```



```{r}
low=(fish$conf.int[1]/(fish$conf.int[1] + 1))*100
hi=(fish$conf.int[2]/(fish$conf.int[2] + 1))*100
```


## The fisher exact test is an odds-ratio test. It is commonly used in clinical trials to test the effectiveness of preventitive drugs like vaccines. In this case it says the odds of getting covid is `r fish$estimate` times greater if you had taken the placebo than if you had taken the vaccine. The media is reporting this as 92% effective. It also gives you a confidence interval that has the true odds ratio within it 95% of the time.  That interval is `r fish$conf.int[1]` to `r fish$conf.int[2]`


## Conclusion effectiveness is simply turning odds (in this case an odds_ratio) into a percentage If the odds of something happening is O, then the probability of it happening is O/(1 + O). To turn probability to a percent just multiply by 100. However effectivness should not just be a number - It should be cast as a confidence interval.



## Effectiveness Confidence Interval `r low`  to `r hi`

